---
title: "506Assignment5"
author: "Jiaqi Sun"
format: 
  html:
    embed-resources: true
editor: visual
---

Github repository: [Github](https://github.com/AlanGalaxy/506A5)

## Problem 1

Load the package and data.

```{r}
library(ggplot2)
library(tidyverse)

nnmaps <- read.csv("D:/Code/R/506A5/chicago-nmmaps.csv")
nnmaps$date <- as.Date(nnmaps$date)
# plot(nnmaps$o3 ~ nnmaps$date)
summary(nnmaps)
```

\(a\)

Calculate the average temperature in Celsius. Use `ggplot` to plot the data.

```{r}
#| code-fold: true
month_mean <- nnmaps %>% 
  group_by(month, season) %>% 
  # get the average temp
  summarise(mean_temp = mean(temp)) %>% 
  # change to celsius
  mutate(mean_temp = (mean_temp - 32) / 1.8) %>% 
  ungroup()

ggplot(data = month_mean, 
       mapping = aes(x = month, y = mean_temp,color = season, group = season)) + 
  # plot the points and lines
  geom_point() + 
  geom_line() + 
  # set the legend and order
  scale_color_manual(name = "Seasons", values = c("Winter" = "blue", 
                                                  "Spring" = "green", 
                                                  "Summer" = "orange", 
                                                  "Autumn" = "black"), 
                     limits = c("Winter", "Spring", "Summer", "Autumn")) + 
  # set the x-axis
  scale_x_discrete(limits = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", 
                            "Aug", "Sep", "Oct", "Nov", "Dec")) + 
  # set the labels
  labs(x = "Month", y = "Average Monthly Temperature (Celsius)", color = "Season", 
       title = "Average Monthly Temperature by Season")
```

\(b\)

Calculate the data and pivot into long data. Use `ggplot` to plot the data.

```{r}
#| code-fold: true
library(tidyr)

month_mean2 <- nnmaps %>% 
  group_by(month, season) %>% 
  # calculate the average data
  summarise(mean_temp = mean(temp), 
            mean_o3 = mean(o3), 
            mean_pm10 = mean(pm10, na.rm = TRUE), 
            mean_dewpoint = mean(dewpoint)) %>% 
  # change to celsius
  mutate(mean_temp = (mean_temp - 32) / 1.8) %>% 
  ungroup()%>%
  # pivot the data to be longer
  pivot_longer(cols = starts_with("mean_"), names_to = "variable", 
               values_to = "value")

ggplot(month_mean2, aes(x = month, y = value, color = season, group = interaction(season, variable), linetype = variable, shape = variable)) +
  geom_point() +
  geom_line() +
  # set the legend and seasons' color
  scale_color_manual(name = "Seasons", values = c("Winter" = "blue", 
                                                  "Spring" = "green", 
                                                  "Summer" = "orange", 
                                                  "Autumn" = "red"), 
                     limits = c("Winter", "Spring", 
                                "Summer", "Autumn")) + 
  # set the legend and point type
  scale_shape_manual(name = "Variables", values = c("mean_temp" = 8, 
                                                    "mean_o3" = 17, 
                                                    "mean_pm10" = 15, 
                                                    "mean_dewpoint" = 19), 
                     labels = c("mean_temp" = "Temperature (Celsius)", 
                                "mean_o3" = "O3",
                                "mean_pm10" = "PM10",
                                "mean_dewpoint" = "Dewpoint"), 
                     limits = c("mean_dewpoint", "mean_pm10", 
                                "mean_o3", "mean_temp")) + 
  # set the legend and line type
  scale_linetype_manual(name = "Variables", values = c("mean_temp" = "dotted", 
                                                       "mean_o3" = "dashed", 
                                                       "mean_pm10" = "solid", 
                                                       "mean_dewpoint" = "dotdash"),
                        labels = c("mean_temp" = "Temperature (Celsius)", 
                                   "mean_o3" = "O3",
                                   "mean_pm10" = "PM10",
                                   "mean_dewpoint" = "Dewpoint"), 
                        limits = c("mean_dewpoint", "mean_pm10", 
                                   "mean_o3", "mean_temp")) + 
  # set the x-axis
  scale_x_discrete(limits = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", 
                            "Aug", "Sep", "Oct", "Nov", "Dec")) +
  # set the labels
  labs(x = "Month", y = "Average Monthly Value", color = "Variable",
       title = "Average Monthly Temperature, O3, PM10, and Dewpoint")
```

The data in different seasons are in different colors, and different variables are in different point-type and line-type.

From the picture, "PM10" is the variable has the least seasonal trend.

## Problem 2

\(a\)

Set the class "poly".

```{r}
setClass("poly", 
         slots = c(coef = "numeric", 
                   degree = "numeric"))
```

Define the constructor. The first vector represents the coefficients, the second represents the degree.

```{r}
makePoly <- function(x, y) {
  # first use the validator defined below to check if the two vector make sense
  new("poly", coef = x, degree = y)
  # order the two vector by the degree of x
  index <- order(y, decreasing = TRUE)
  x <- x[index]
  y <- y[index]
  # return the ordered poly
  return(new("poly", coef = x, degree = y))
}
```

Define the validator to check whether the two vectors have the same length.

```{r}
setValidity("poly", function(object) {
  # check the same length
  if (length(object@coef) != length(object@degree)) {
    stop("The coefficients and degrees do not have the same length!\n")
  }
  return(TRUE)
})
```

Define the `show` method.

```{r}
#| code-fold: true
setMethod("show", "poly", function(object) {
  coef <- object@coef
  degree <- object@degree
  equation <- ""
  
  # Print the first term
  # In order to make sure the equation is well presented, this can be separated into 
  # different cases.
  if (degree[1] == 0) {
    if (coef[1] > 0) {
      equation <- as.character(coef[1])
    } else {
      equation <- paste("- ", abs(coef[1]), sep = "")
    }
  } else if (degree[1] == 1) {
    if (coef[1] == 1) {
      equation <- "x"
    } else if (coef[1] > 0) {
      equation <- paste(coef[1], "x", sep = "")
    } else if (coef[1] == -1) {
      equation <- "- x"
    } else {
      equation <- paste("- ", abs(coef[1]), sep = "")
    }
  } else {
    if (coef[1] == 1) {
      equation <- paste("x^", degree[1], sep = "")
    } else if (coef[1] > 0) {
      equation <- paste(coef[1], "x^", degree[1], sep = "")
    } else if (coef[1] == -1) {
      equation <- paste("- ", "x^", degree[1], sep = "")
    } else {
      equation <- paste("- ", abs(coef[1]), "x^", degree[1], sep = "")
    }
  }
  
  # Print the rest terms
  if (length(coef) > 1) {
    for (i in 2:length(coef)) {
      if (degree[i] == 0) {
        if (coef[i] > 0) {
          equation <- paste(equation, " + ", coef[i], sep = "")
        } else {
          equation <- paste(equation, " - ", abs(coef[i]), sep = "")
        }
      } else if (degree[i] == 1) {
        if (coef[i] == 1) {
          equation <- paste(equation, " + x", sep = "")
        } else if (coef[i] > 0) {
          equation <- paste(equation, " + ", coef[i], "x", sep = "")
        } else if (coef[i] == -1) {
          equation <- paste(equation, " - x", sep = "")
        } else {
          equation <- paste(equation, " - ", abs(coef[i]), "x", sep = "")
        }
      } else {
        if (coef[i] == 1) {
          equation <- paste(equation, " + x^", degree[i], sep = "")
        } else if (coef[i] > 0) {
          equation <- paste(equation, " + ", coef[i], "x^", degree[i], sep = "")
        } else if (coef[i] == -1){
          equation <- paste(equation, " - x^", degree[i], sep = "")
        } else {
          equation <- paste(equation, " - ", abs(coef[i]), "x^", degree[i], sep = "")
        }
      }
    }
  }
  cat(equation, "\n")
  return(invisible(object))
})
```

Define the addition & subtraction.

```{r}
#| code-fold: true
setMethod("+", signature("poly", "poly"), 
          function(e1, e2) {
            # get all the degrees in two poly
            all_degrees <- unique(c(e1@degree, e2@degree))
            # add the corresponding coefficients
            result_coefficients <- sapply(all_degrees, function(power) {
              sum(e1@coef[e1@degree == power], na.rm = TRUE) + 
                sum(e2@coef[e2@degree == power], na.rm = TRUE)
            })
            # return the new poly
            makePoly(result_coefficients, all_degrees)
          })

setMethod("-", signature("poly", 
                         "poly"), 
          function(e1, e2) {
            all_degrees <- unique(c(e1@degree, e2@degree))
            result_coefficients <- sapply(all_degrees, function(power) {
              sum(e1@coef[e1@degree == power], na.rm = TRUE) - 
                sum(e2@coef[e2@degree == power], na.rm = TRUE)
            })
            
            makePoly(result_coefficients, all_degrees)
          })
```

\(b\)

```{r}
p1 <- makePoly(c(3,2), c(2,0))
p2 <- makePoly(c(7,-2,-1,17), c(3,2,1,0))

p1
p2

p1 + p2
p1 - p2
```

As demonstrated above, $x^1$ and $x^0$ are well handled.

Additional examples:

$\pm1$ in coefficient is well handled.

```{r}
makePoly(c(1,-1), c(2,1))
```

Negative degree is well handled.

```{r}
makePoly(c(-1,2,3), c(0,-1,-2))
```

Non-integer is well handled.

```{r}
makePoly(c(-1,2,3), c(1.5, 0.5,-0.5))
```

Unordered input in degree is well handled.

```{r}
makePoly(c(1,2,3), c(1,0,2))
```

## Problem 3

Load the package and data.

```{r}
library(data.table)
library(nycflights13)

data(flights)
data(airports)

flights <- as.data.table(flights)
airports <- as.data.table(airports)
```

\(a\)

Generate the full `data.table` for departure delay.

```{r}
departure_delay <- flights |> 
  # calcualte the mean and median statistics
  _[, .(mean_dep_delay = mean(dep_delay, na.rm = TRUE), 
        median_dep_delay = median(dep_delay, na.rm = TRUE)), 
    by = .(origin)] |> 
  # merge two data.table
  merge(x = _, airports, by.x = "origin", by.y = "faa", all.x = TRUE) |> 
  # select the final table
  _[, .(airport_name = name, mean_dep_delay, median_dep_delay)] |> 
  # make it in right order
  _[order(-mean_dep_delay)]

print(departure_delay, n = nrow(departure_delay))
```

Generate the full `data.table` for arrival delay.

```{r}
arrival_delay <- flights |> 
  # calcualte the mean and median statistics
  _[, .(mean_arr_delay = mean(arr_delay, na.rm = TRUE), 
        median_arr_delay = median(arr_delay, na.rm = TRUE),
        count = .N), 
    by = .(dest)] |> 
  # merge two data.table
  merge(x = _, airports, by.x = "dest", by.y = "faa", all.x = TRUE) |> 
  # select the final table
  _[count >= 10, .(airport_name = name, mean_arr_delay, median_arr_delay)] |> 
  # make it in right order
  _[order(-mean_arr_delay)]

print(arrival_delay, n = nrow(arrival_delay))
```

\(b\)

Generate the full `data.table`.

```{r}
data(planes)
planes <- as.data.table(planes)

fast_model_num <- flights |> 
  # merge two data.table
  merge(x = _, planes, by = "tailnum", all.x = TRUE) |> 
  # calculate the average speed in MPH for each model
  _[, .(average_speed = mean(distance / air_time, na.rm = TRUE) * 60, 
        num_flight = .N), by = model] |> 
  # make it in right order
  _[order(-average_speed)] |> 
  # select the first row
  _[1, .(model, average_speed, num_flight)]

print(fast_model_num, n = nrow(fast_model_num))
```
